---
title: "Assignment 12"
author: "Ganesh Kodi"
date: "November 20, 2018"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com")) # Setting Cran options
install.packages("dygraphs")
install.packages("fpp2")
install.packages("forecast")
install.packages("xts")
library("dygraphs")
library(forecast)
library(fpp2)
library(xts)


```

### Warm Up: Brief Financial Data 
#### Summary
EuStockMarkets is sample data sets of prominent stocks over time avaialale natively in R. Extract DAX index data from European stock data (EuStockMarkets) and analyse the DAX data. Firstly, checking whether the data is Additive or Mulitplicative data and later, decide whether transformation required or not

```{r EuStockMarkets Summary, echo=TRUE}
#Display EuStockMarkets dataset
summary(EuStockMarkets)
#Extract DAX timeseries dataset from EuStockMarkets
dax <- EuStockMarkets[,1]
summary(dax) # Summary of the DAX index
```
##### Inference: 
Based on summary statistics, maximum vs minimum value are not huge ratio i.e. more than 10, Transformation is not necessary. Model is of Additive type. Continue with original DAX data would be prefered for analysis. 

#### Plot 
Rudimentary plot for European stock data and DAX stock data
```{r EuStockMarkets Plot, echo=TRUE}
# Plot the line graph of the European Stock Markets Data
plot(EuStockMarkets, col="blue", xlab="Year", ylab="Index value", main="European stock data analysis of 90's" )
abline(v=1997, col="red")
# Plot the line graph of the DAX Stock Markets Data
plot(dax, col="blue", xlab="Year", ylab="Index value", main="DAX stock data analysis of 90's" )
abline(v=1997, col="red")
```

#### Decompose DAX stock data 
Decompose DAX stock data into Trend, Seasonal and Random components. Make it multiplicative model
```{r DAX Decompose, echo=TRUE}
daxcomp <- decompose(dax, type = "mult")
plot(daxcomp, col="blue")
abline(v=1997, col="red")
```

### 2.	Temperature Data 
#### Summary
Maxtemp is dataset contains time series data of Maximum annual temperatures (degrees Celsius) for Moorabbin Airport, Melbourne. 1971-2016.
```{r maxtemp, echo=TRUE}
autoplot(maxtemp) # Command in the example section of help of maxtemp

```

#### Subset
Subsetting the data after 1990
```{r maxtemp subset, echo=TRUE}
subtemp <- subset(maxtemp, start=20) #Subsetting time series data from 1990 which is 20th index

```

Using SES predict next five years of maximum temperature of Melbourne. Plot forecast and predicted line overlay on the plot
```{r ses with subtemp, echo=TRUE }
fitsubtemp <- ses(subtemp, h=5) # Creating SES linear trend of the subset
plot(fitsubtemp) # Plot the fit
lines(fitted(fitsubtemp),col="blue") # Overlay the predicated line over plot
summary(fitsubtemp) # Using summary collect AICs value
```

Using holt linear method predict next five years of maximum temperature of Melbourne. Plot forecast and predicted line overlay on the plot
```{r holt with subtemp, echo=TRUE }
fitholtsubtemp <- holt(subtemp, initial="optimal", h=5) # Creating holt linear tred of the subset
plot(fitholtsubtemp) # Plot the fit
lines(fitted(fitholtsubtemp),col="blue") # Overlay the predicated line over plot
summary(fitholtsubtemp) # Using summary collect AICs value
```
##### AICc for the different model are
####### 1. SES  --> 141.53
####### 2. Holt --> 142.23

### 3.	The Wands Choose the Wizard 
#### Summary
Creating the timeseries of Wands from dataset provided and convert it to xts object and Using dygraph package analyse the timeseries
```{r wands summary, echo=TRUE}
df_gregowitch <- read.csv("source/Unit11TimeSeries_Gregorovitch.csv") #Creating dataset from Unit11TimeSeries_Gregorovitch.csv in source folder 
df_ollivander <- read.csv("source/Unit11TimeSeries_Ollivander.csv") #Creating dataset from Unit11TimeSeries_Ollivander.csv in source folder
names(df_gregowitch) <- c("Year", "Numbers") #Name the column headers
names(df_ollivander) <- c("Year", "Numbers") #Name the column headers
df_gregowitch$Year <- as.Date.character(df_gregowitch$Year,format = "%d/%m/%Y") #Format the Year column proper date format
df_ollivander$Year <- as.Date.character(df_ollivander$Year,format = "%d/%m/%Y") #Format the Year column proper date format
xts_gregowitch <- as.xts(read.zoo(df_gregowitch)) #Convert data frame to xts object
xts_ollivander <- as.xts(read.zoo(df_ollivander)) #Convert data frame to xts object
merge_wands <-merge(xts_gregowitch,xts_ollivander) #Combine both the xts object
dygraph(merge_wands,main = "Wand sales comparison between wand makers", xlab = "Year", ylab = "Numbers") %>%   #Create dygraph
dySeries("xts_ollivander", label= "Olivander",drawPoints = TRUE,  color = "blue") %>%  dySeries("xts_gregowitch", label="Gregowitch", drawPoints = TRUE , color = "green") %>%  #Label and color the series
  dyRangeSelector() %>% #Activate range selector
dyShading(from = "1995-1-1", to = "1999-1-1", color = "orange") #Sharding the time between 1995-1999



```

Git URL for this assignment is as follows
https://github.com/ganeshkodi/smumsds_homework/tree/master/Homework12 
